%!TEX root = ../Project.tex

\subsection{Pilot by the Archaeology and History departments}
\label{sec:developmentpilot}

During the spring term (January to March 2012), the Archaeology and History
departments at the University of York used the software developed for this
project to allocate modules to their second and third-year students. The
application documentation (Appendix~\ref{sec:documentation}) includes
information for departmental administrators about the setup of the application
and limitations they should be aware of.

A production instance of the application \gls{vm} was provided by
\gls{itservices} on 22 February, and the application was promoted from
\texttt{wwwtest.york.ac.uk} to \texttt{www.york.ac.uk} later that day after
final sign-off from the University's Information Security Officer. The
application was successfully set up by departmental administrators on 23
February and the setup was confirmed by departments to the \gls{aso} after a
project group meeting that day.
Table~\ref{development_pilot_department_numbers} gives the amount of
information added to the system by the pilot departments.

\begin{table}
  \begin{center}
    \begin{tabular}{ | l | l | l | l | }
      \hline
      \textbf{Department}  & \textbf{Types of student (by degree)} & \textbf{Modules} & \textbf{Students} \\
      \hline
      History     & 14                           & 57      & 591      \\
      Archaeology & 8                            & 32      & 204      \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Amount of data loaded into the application by the pilot departments}
  \label{development_pilot_department_numbers}
\end{table}

As the application was hosted by \gls{itservices} using the
\texttt{york.ac.uk} domain name, notification was given to the team that
provides end-user support to students and staff. Issues are logged by this
team from a variety of sources, including emails to
\texttt{itsupport@york.ac.uk}, telephone calls and visits to the IT Support
office on campus. During the trial period any issues relating to this
application were passed to departmental administrators and the
\gls{itservices} developer assisting with the application.

The application was available to students from 0900 on Monday 27 February
until 1700 on Monday 5 March. Only one issue directly relating to the
application code was raised by a user to their department during the week. The
user reported they were unable to use the drag-and-drop functionality to make
their choices. Support details provided by the user revealed that they were
using Internet Explorer 7, a web browser that was superseded by Internet
Explorer 8 in March 2009. As of the writing of this report, computers provided
on campus at the University of York run IE8, as well as alternate web browsers
such as Firefox and Opera (all of which were used to test while creating this
application).

The project implementer had intended to include some text for users of old
browsers, indicating that the application had not been tested for them and
they should use an alternate browser if possible. Sites such as
\url{http://www.ie6nomore.com/} include snippets of code that demonstrate how
to achieve this. Unfortunately, for some reason this code was not committed to
the repository and did not make it into the release version of the software.
The student was advised by the departmental administrator to use a more
up-to-date web browser if possible and they successfully used a
University-provided computer to select their modules within 24 hours of the
problem initially being raised.

Several students who were on study aboard programs reported issues connecting
to the University's \gls{vpn}, though these issues are out of scope for this
project. The issues could not be remedied quickly by \gls{itservices} due to
the lack of detailed diagnostic information available from the students
encountering issues, so they simply sent their choices by email. The choices
were input into the system by administrators acting on behalf of students. It
is worth noting here that if this were an application built and supported by
the University (as it will be if it is available in future years), it would be
available away from campus and these \gls{vpn} issues would not have arisen.

On the afternoon of Friday 2 March the software was upgraded to include
improvements made during the week. Clearly performing a software upgrade the
afternoon before a weekend is not ideal -- if there had been any issues, they
could not have been fixed quickly. However, the \gls{itservices} developer
responsible for helping with the application was away from the office for
three days after the weekend. I decided that the upgrade had to be performed
before then in order for departments to attempt to make an allocation sooner
rather than later. If any issues had arisen during the upgrade, the developer
would have reverted to the previous working release version.

On the morning of Monday 5 March the History department reported they had
received emails from approximately ten students who had encountered login
errors when trying to use the application. \gls{itservices} noted that the
University had experienced problems with its authentication systems
(Shibboleth) over the weekend. Departments were then informed this was not an
issue with the module allocation application and should at that point be
resolved. Some History students had their choices input by the department as a
result of this issue.

A History administrator stated in an email that by 10am on Monday 5 March
(96\% of the way through the application's opening time) just 71\% of students
had used the system to make their choices. However the department was
unconcerned, commenting that with time-limited processes such as this they
found that students tended to wait until just before the deadline.
Departmental administrators commented that they thought students had ``engaged
well'' with the software during the pilot week. Several students emailed their
departments to recommend an enhancement to the interface. They requested that
when making a choice for the second time, the ranking interface
(Figure~\ref{walkthrough_student_rank}) should not look as though it has never
been used before -- it should load the choices made the previous time. This
was not discovered during user testing as students were never given the
opportunity to use the system more than once.

In all, only one student raised an issue that prevented them using the
application during the week, which is equivalent to approximately 0.13\% of
all users. I would therefore declare this pilot to be a fairly successful
trial of the student interface.

The History department performed an allocation on the evening of 5 March,
which completed with Gurobi reporting an ``optimal'' status. At approximately
midday on 6 March the system stopped working, giving an ``Application Error''
to end users (this means an exception is being thrown in the Java code). On
investigation with \gls{itservices}, the database contained one less module
availability than module (as mentioned in
Section~\ref{developmentdatabasestructure}, these are separate entities). This
removal was either caused by a bug in the application code that was impossible
to reproduce during testing and has not since reappeared, or by user error
that has not been discovered (accidental deletion) -- unfortunately, it is
impossible to definitively state the cause. The removal of that entity
resulted in the removal of all the student ranks associated with that module,
and so the database had to be restored to the nightly backup made by
\gls{itservices} on 6 March at 3am. This restoration occurred on 9 March and
the History department were able to perform another allocation. From Monday 12
March the department spent several days modifying class sizes and manually
reallocating several students to improve their allocation. This was a
time-consuming process as it involved decisions being made by the Chair of the
Board of Studies about staffing of teaching posts for the next academic year.

The Archaeology department performed an allocation on 6 March. The allocation
resulted in one module not being run (the solver found an optimal allocation
without using that module), and the decision to discontinue the module had to
be referred to the department's Board of Studies (chaired by \gls{sa}). The
allocation of students to modules created by Gurobi was described as ``a
triumph'' by the department, and they were able to pass a file containing the
allocations to \gls{sits} almost one week before the University deadline.

\Gls{mh} imported (into \gls{sits}) the Archaeology data on 12 March and the
History data on 16 March. He commented that the Archaeology data had some
discrepancies caused by some students changing degree type while the
application was live, but that History had provided ``a cleaner data set''. He
mentioned the Archaeology data had taken him approximately one hour to import,
which is far less than the several hours it would have taken the Archaeology
department to input the data student-by-student.
