%!TEX root = ../Project.tex

\subsection{Implementation schedule}

The project commenced on 13 May 2011 with the first meeting of the project
steering group. User research and the literature review took place in the
autumn term, while the spring term was set aside solely for implementation.

As the time allotted for implementation of the software was limited to under
eight weeks, it was necessary to prioritise certain aspects of the system in
case I was unable to implement every requirement.

Fortunately, the various elements of this project are naturally quite loosely
coupled. I identified the following three aspects (ordered by date required):

\begin{enumerate}
  \item Student interface to collect data (27 Feb 2012)
  \item Performing the allocation (5 March 2012)
  \item Staff interface (approx. 16 March 2012)
\end{enumerate}

The staff interface is given as the 16 March as it is not critical for this
pilot. While it would clearly be preferable for staff to set up the system
themselves unaided, it is more important that there is a working student
interface and allocation algorithm.

\subsection{Gantt chart}

A Gantt chart (given in Figure~\ref{ganttchart}) was created to show the
expected progress for the project's implementation. Each of the blocks refers
to week numbers at the University of York, where Autumn week 4 begins on the
31 October 2011, and Spring week 1 on the 9 January 2012.

This Gantt chart shows the conflict in timelines between the project as a
final-year Computer Science project and as a University software project with
deliverables. The software was live and available to students and staff at the
end of the spring term, the busiest time for writing this report.

\input{inc/gantt.tex}
